---
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
mathjax: null
params:
  directory_output_reports: "tomt"
  directory_and_filename_function_file: "tomt"
  projectname: "tomt"
  date_of_analysis: 0
  directory_input_matrix_sampleID: "tomt"
  filename_matrix: "tomt"
  filename_sampleID: "tomt"
  no_permutations_over_vs: 0
  no_permutations_post_vs_selected_models: 0
  filter_percent_in_each_group: 0
  groupsnumeric: "tomt"
  model_strategies_to_run: 0
  max_no_of_ortho_pre_vs: 0
  max_no_of_ortho_post_vs: 0
  prefered_pR2_and_pQ2_permutated_post_vs: 0
  reordered_levels_of_groups: 0
  pcorr_diff: 0
  variable_selection_using_VIP: "tomt"

---

---
title: "Summary ropls models `r gsub("_"," ",paste(params$projectname))`"
author: |
    | roplspvs - Ropls with permutated variable selection: Version 0.15.0
    | Marika Strom
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, echo=FALSE}

directory_output_reports <- params$directory_output_reports
directory_and_filename_function_file <- params$directory_and_filename_function_file
projectname<-params$projectname
date_of_analysis<-params$date_of_analysis
directory_input_matrix_sampleID<-params$directory_input_matrix_sampleID
filename_matrix<-params$filename_matrix
filename_sampleID<-params$filename_sampleID
no_permutations_over_vs<-params$no_permutations_over_vs
no_permutations_post_vs_selected_models<-params$no_permutations_post_vs_selected_models
filter_percent_in_each_group<-params$filter_percent_in_each_group
groupsnumeric<-params$groupsnumeric
model_strategies_to_run <- params$model_strategies_to_run
max_no_of_ortho_pre_vs <- params$max_no_of_ortho_pre_vs
max_no_of_ortho_post_vs <- params$max_no_of_ortho_post_vs
prefered_pR2_and_pQ2_permutated_post_vs <- params$prefered_pR2_and_pQ2_permutated_post_vs
reordered_levels_of_groups <- params$reordered_levels_of_groups
pcorr_diff <- params$pcorr_diff
variable_selection_using_VIP <- params$variable_selection_using_VIP


knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE) 
knitr::opts_knit$set(root.dir=paste(directory_output_reports))

library("ropls")
library(kableExtra)
library(gridExtra)
library("ggpubr")
library("matrixStats")
library(stringr)
library(devtools)
library(DescTools)

source(paste(directory_and_filename_function_file))
directory_Rmarkdownfiles <- paste(directory_of_roplspvs,"/",foldername_Rmarkdownfiles,"/",sep="")

```


Project name: `r projectname`

Date of analysis: `r as.character(date_of_analysis)`

Directory data matrix and sample id: `r directory_input_matrix_sampleID`

Filename data matrix: `r filename_matrix`

Filname sample id: `r filename_sampleID`

Directory and file name function file: `r directory_and_filename_function_file`

Directory rdata files: `r directory_output_reports`

Permutations over variable selection: `r no_permutations_over_vs`

Permutations post variable selection in selected models: `r no_permutations_post_vs_selected_models`

Missing value tolerance in each group: `r filter_percent_in_each_group`

order of groups with diseased first followed by controls: `r reordered_levels_of_groups`

max number of orthogonal variables in model pre variable selection: `r max_no_of_ortho_pre_vs`

max number of orthogonal variables in model post variable selection: `r max_no_of_ortho_post_vs`

prefered pR2 and pQ2 during model selection permutated post variable selection: `r prefered_pR2_and_pQ2_permutated_post_vs`

pcorr increase allowing Q2 to decrease 1% during model selection in Model4: `r pcorr_diff`

variable selection method: `r if(variable_selection_using_VIP=="yes") {paste("Using p(corr) and VIP")} else {paste("Using p(corr)")}`

```{r, model strategy 1}

if (1 %in% model_strategies_to_run){
  resultmodelname<-"result_Model1"
  permutated_modelsname<-"permutated_models_Model1"
  percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_name<- "percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_Model1"
description_Model1 <- "Models with userset p(corr) cutoff during variable selection and userset number of orthogonal variables pre and post variable selection"
model_description <- description_Model1
model_strategy<-1}
  
```

```{r, child=if (1 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_each_strategy.Rmd',sep="")}

```
```{r, model strategy 2}
if (2 %in% model_strategies_to_run){
  resultmodelname<-"result_Model2"
  permutated_modelsname<-"permutated_models_Model2"
  percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_name<- "percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_Model2"
description_Model2 <- "Model with p(corr) cutoff corresponding to userset pvalue cutoff for correlation"
model_description  <- description_Model2
model_strategy<-2}
  
```

```{r, child=if (2 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_each_strategy.Rmd',sep="")}

```
```{r, model strategy 3}
if (3 %in% model_strategies_to_run){
  resultmodelname<-"result_Model3"
  permutated_modelsname<-"permutated_models_Model3"
  percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_name<- "percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_Model3"
description_Model3 <- "Model with variable using p(corr) cutoff resulting in best performing model post variable selection"
model_description  <- description_Model3
model_strategy<-3}
  
```

```{r, child=if (3 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_each_strategy.Rmd',sep="")}

```

```{r, model strategy 4}
if (4 %in% model_strategies_to_run){
  resultmodelname<-"result_Model4"
  permutated_modelsname<-"permutated_models_Model4"
  percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_name<- "percent_R2_and_Q2_in_permutated_larger_than_in_unpermutated_Model4"
description_Model4 <- "Model with variable using p(corr) cutoff resulting in best performing model with userset max p(pearson)  and using minimal amount of variables"
model_description  <- description_Model4
model_strategy<-4}
  
```

```{r, child=if (4 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_each_strategy.Rmd',sep="")}

```

```{r, model strategy 5}
if (5 %in% model_strategies_to_run){
  resultmodelname<-"result_iterationmodel"
description_Model5<- "Iteration models removing variables with lowest p(corr) iteratively"
model_description <- description_Model5
model_strategy<-5}
  
```

```{r, child=if (5 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_iteration_model.Rmd',sep="")}

```

```{r, pcorr tables model strategy 2}
if (2 %in% model_strategies_to_run){
  resultmodelname<-"result_Model2"
 model_description <- description_Model2
model_strategy<-2}
  
```

```{r, child=if (2 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_pcorr_selected_variables.Rmd',sep="")}

```

```{r, pcorr tables model strategy 3}
if (3 %in% model_strategies_to_run){
  resultmodelname<-"result_Model3"
 model_description <- description_Model3
model_strategy<-3}
  
```

```{r, child=if (3 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_pcorr_selected_variables.Rmd',sep="")}

```

```{r, pcorr tables model strategy 4}
if (4 %in% model_strategies_to_run){
  resultmodelname<-"result_Model4"
 model_description <- description_Model4
model_strategy<-4}
  
```

```{r, child=if (4 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_pcorr_selected_variables.Rmd',sep="")}

```

```{r, pcorr tables model strategy 5}
if (5 %in% model_strategies_to_run){
  resultmodelname<-"result_iterationmodel"
 model_description <- description_Model5
model_strategy<-5}
  
```

```{r, child=if (5 %in% model_strategies_to_run) paste(directory_Rmarkdownfiles,'/roplspvs_Summary_pcorr_selected_variables.Rmd',sep="")}

```

```{R}

save(list = ls(all.names = TRUE), file = paste(directory_output_reports,"/", "Summary_",projectname, "_",date_of_analysis,".Rdata", sep=""), envir = environment())

```

