---
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
mathjax: null
params:
  group1: "tomt"
  group2: "tomt"
  secID: "tomt"
  pcorr_Model1: 0
  ortho_pre_vs_Model1: 0
  ortho_post_vs_Model1: 0
  setseedno: 0
  p_pearson_of_pcorr_cutoff: 0
  directory_input_matrix_sampleID: "tomt"
  filename_matrix: "tomt"
  decimal_separator: "tomt"
  variable_names_length: 0
  variable_names_position: "tomt"
  filename_sampleID: "tomt"
  directory_output_reports: "tomt"
  projectname: "tomt"
  date_of_analysis: 0
  no_permutations_post_vs: 0
  no_permutations_post_vs_selected_models: 0
  no_permutations_over_vs: 0
  colname_groupID: "tomt"
  colname_secID: "tomt"
  replace_0: "tomt"
  filter_percent_in_each_group: 0
  replace_NA: "tomt"
  log_transform: "tomt"
  directory_and_filename_function_file: "tomt"
  max_no_of_ortho_pre_vs: 0
  max_no_of_ortho_post_vs: 0
  prefered_pR2_and_pQ2_permutated_post_vs: 0
  reordered_levels_of_groups: 0
  pcorr_diff: 0
  variable_selection_using_VIP: "tomt"
  model_strategies_to_run: 0

---

---
title: |
    | ROPLS models `r gsub("_"," ",paste(params$projectname))`
    | Model of group `r group1` versus `r group2` for `r secID`
 
author: |
    | Ropls with permutated variable selection: Version 0.15.0
    | Marika Strom
date: `r Sys.Date()`
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = TRUE, fig.fullwidth = TRUE)
# opts_chunk$set sets the parameters for the whole document
```

```{r, echo=FALSE}

library("ropls")
library("ggplot2")
library("ggrepel")
library(kableExtra)
library(gridExtra)
library("ggpubr")
library("matrixStats")
library(stringr)
library(tryCatchLog)
library(DescTools)
library(precrec)
library(pROC)
library(rstatix)
```

```{r echo=FALSE}

group1 <- params$group1
group2 <- params$group2
secID <- params$secID
pcorr_Model1 <- params$pcorr_Model1
ortho_pre_vs_Model1 <- params$ortho_pre_vs_Model1
ortho_post_vs_Model1 <- params$ortho_post_vs_Model1
setseedno <- params$setseedno
p_pearson_of_pcorr_cutoff <- params$p_pearson_of_pcorr_cutoff
directory_input_matrix_sampleID <- params$directory_input_matrix_sampleID
filename_matrix <- params$filename_matrix
decimal_separator <- params$decimal_separator
variable_names_length <- params$variable_names_length
variable_names_position <- params$variable_names_position
filename_sampleID <- params$filename_sampleID
directory_output_reports <- params$directory_output_reports
projectname <- params$projectname
date_of_analysis <- params$date_of_analysis
no_permutations_post_vs <- params$no_permutations_post_vs
no_permutations_post_vs_selected_models <- params$no_permutations_post_vs_selected_models
no_permutations_over_vs <- params$no_permutations_over_vs
colname_groupID <- params$colname_groupID
colname_secID <- params$colname_secID
replace_0 <- params$replace_0
filter_percent_in_each_group <- params$filter_percent_in_each_group
replace_NA <- params$replace_NA
log_transform <- params$log_transform
directory_and_filename_function_file <- params$directory_and_filename_function_file
max_no_of_ortho_pre_vs <- params$max_no_of_ortho_pre_vs
max_no_of_ortho_post_vs <- params$max_no_of_ortho_post_vs
prefered_pR2_and_pQ2_permutated_post_vs <- params$prefered_pR2_and_pQ2_permutated_post_vs
reordered_levels_of_groups <- params$reordered_levels_of_groups
pcorr_diff <- params$pcorr_diff
variable_selection_using_VIP <- params$variable_selection_using_VIP
model_strategies_to_run <- params$model_strategies_to_run

directory_and_filename_rdata <- paste(paste(directory_output_reports, projectname, sep = ""), date_of_analysis, group1, "vs", group2, secID, ".Rdata", sep = "_")
filename_rdata <- paste(projectname, date_of_analysis, group1, "vs", group2, secID, ".Rdata", sep = "_")
source(directory_and_filename_function_file)
```

## Data to analysis

```{R loading data}
# Loading matrix and metadata

directory_and_filename_input_matrix <- paste(directory_input_matrix_sampleID, "/", filename_matrix, sep = "")
directory_and_filename_sampleID <- paste(directory_input_matrix_sampleID, "/", filename_sampleID, sep = "")

# read datamatrix
if (decimal_separator == "dot") {
  datamatrix <- read.csv(directory_and_filename_input_matrix, header = T, dec = ".", row.names = 1, check.names = FALSE, na.strings = c("", "NA", "Inf"), sep = "\t")
} else {
  datamatrix <- read.csv(directory_and_filename_input_matrix, header = T, dec = ",", row.names = 1, check.names = FALSE, na.strings = c("", "NA", "Inf"), sep = "\t")
}

# Read meta data with sample information for annotation####
if (decimal_separator == "dot") {
  sampleID <- read.table(directory_and_filename_sampleID, header = T, dec = ".", row.names = 1, check.names = FALSE, na.strings = c("", "NA", "Inf"), sep = "\t")
} else {
  sampleID <- read.table(directory_and_filename_sampleID, header = T, dec = ",", row.names = 1, check.names = FALSE, na.strings = c("", "NA", "Inf"), sep = "\t")
}
```

Project name: `r projectname`

Date of analysis: `r as.character(date_of_analysis)`

Directory data matrix and sample id: `r directory_input_matrix_sampleID`

Filename data matrix: `r filename_matrix`

Filname sample id: `r filename_sampleID`

Directory and file name function file: `r directory_and_filename_function_file`

Directory output reports: `r directory_output_reports`

File name rdata: `r filename_rdata`

Permutations including variable selection in selected models: `r no_permutations_over_vs`

Permutations after variable selection during model selection: `r no_permutations_post_vs`

Permutations after variable selection in selected models: `r no_permutations_post_vs_selected_models`

Missing value tolerance in each group: `r filter_percent_in_each_group`

Model: group `r group1` vs group `r group2` for `r secID`

max number of orthogonal variables in model pre variable selection: `r max_no_of_ortho_pre_vs`

max number of orthogonal variables in model post variable selection: `r max_no_of_ortho_post_vs`

prefered pR2 and pQ2 during model selection permutated post variable selection: `r prefered_pR2_and_pQ2_permutated_post_vs`

pcorr increase allowing Q2 to decrease 1% during model selection in Model strategy 4: `r pcorr_diff`

variable selection method: `r if(variable_selection_using_VIP=="yes") {paste("Using p(corr) and VIP")} else {paste("Using p(corr)")}`

```{R subset sampleID and matrix, error=FALSE}

# subset sampleID and matrix

subsetdatamatrix <- subsetmatrixfunction(sampleID, datamatrix, group1, group2, secID)
subsetsampleID <- subsetsampleIDfunction(sampleID, group1, group2, secID)
class <- as.factor(subsetsampleID[, paste(colname_groupID)])
classordered <- class
inv_reordered_levels_of_groups <- rev(reordered_levels_of_groups)
levels(classordered) <- inv_reordered_levels_of_groups
for (i in 1:length(class)) {
  classordered[i] <- class[i]
}
classordered <- droplevels(classordered)
llq <- min(datamatrix[datamatrix!=0&!is.na(datamatrix)])
lld <- llq/3
if (replace_0!=F) {if (replace_0=="NA") {subsetdatamatrix[subsetdatamatrix==0]<-NA} else if (replace_0=="lld") {subsetdatamatrix[subsetdatamatrix==0]<-lld} else {subsetdatamatrix[subsetdatamatrix==0]<-replace_0}}
subsetdatamatrixfiltered <- filterNAfunction(subsetsampleID, subsetdatamatrix, group1, group2, secID, filter_percent_in_each_group)
if (replace_NA!=F) {if (replace_NA=="lld") {subsetdatamatrixfiltered[is.na(subsetdatamatrixfiltered)]<-lld} else {subsetdatamatrixfiltered[is.na(subsetdatamatrixfiltered)] <- replace_NA}}
if (log_transform==T) {subsetdatamatrixfiltered <- log(subsetdatamatrixfiltered)}
subsetdatamatrixfiltereddf <- as.data.frame(subsetdatamatrixfiltered)
subsetdatamatrixdummies <- subsetdatamatrixfiltereddf
tryCatch(
  {
    subsetdatamatrixdummies <- fastDummies::dummy_cols(subsetdatamatrixfiltereddf, remove_selected_columns = TRUE, ignore_na = TRUE)
    rownames(subsetdatamatrixdummies) <- rownames(subsetdatamatrixfiltereddf)
  },
  error = function(e) {
    rownames(subsetdatamatrixdummies) <- rownames(subsetdatamatrixfiltereddf)
  }
)

subsetdatamatrix <- subsetdatamatrixdummies
for (i in 1:ncol(subsetdatamatrix)) {
  subsetdatamatrix[, i] <- as.numeric(subsetdatamatrix[, i])
}

tsubsetdatamatrix <- t(subsetdatamatrix)
set.seed(setseedno)
```

subject number group `r group1`: `r ngroup1 <- nrow(subset(subsetsampleID,subsetsampleID[,paste(colname_groupID)]==group1)); ngroup1`

subject number group `r group2`: `r ngroup2 <- nrow(subset(subsetsampleID,subsetsampleID[,paste(colname_groupID)]==group2)); ngroup2`

no of variables before filtering: `r no_of_variables_before_filtering <- ncol(datamatrix); no_of_variables_before_filtering`

no of variables after filtering: `r no_of_variables_after_filtering <- ncol(subsetdatamatrixfiltereddf); no_of_variables_after_filtering`

no of expanded variables after filtering: `r no_of_expanded_variables_after_filtering <- ncol(subsetdatamatrixdummies); no_of_expanded_variables_after_filtering`


   

```{R}

if (pcorr_Model1 == "according to p-value") {
  pcorr <- calculatepcorrfrompvalue(selectpvalue = p_pearson_of_pcorr_cutoff, n = nrow(subsetsampleID))
} else {
  pcorr <- pcorr_Model1
}
```
## PCA 
### All subjects included in comparison
```{R PCA}

plot(opls(subsetdatamatrix, typeVc = "x-score", info.txtC = "none", fig.pdfC = "none"), parAsColFcVn = classordered)
```

### Subjects in `r group1`

```{R PCA group1}
subsetdatamatrix_group1 <- subset(subsetdatamatrix, subsetsampleID[, paste(colname_groupID)] == group1)
if (nrow(subsetdatamatrix_group1 < 7)) {
  set_crossvalI <- nrow(subsetdatamatrix_group1)
} else {
  set_crossvalI <- 7
}
opls(subsetdatamatrix_group1, typeVc = "x-score", info.txtC = "none", crossvalI = set_crossvalI)
```

### Subjects in `r group2`

```{R PCA group2}
subsetdatamatrix_group2 <- subset(subsetdatamatrix, subsetsampleID[, paste(colname_groupID)] == group2)
if (nrow(subsetdatamatrix_group2 < 7)) {
  set_crossvalI <- nrow(subsetdatamatrix_group2)
} else {
  set_crossvalI <- 7
}
opls(subsetdatamatrix_group2, typeVc = "x-score", info.txtC = "none", crossvalI = set_crossvalI)
```


```{r, child=if (1 %in% model_strategies_to_run) 'Ropls_with_permutated_variable_selection_Models_of_each_comparison_strategy1.Rmd'}

```

```{r}
first_analysis<-T
if(file.exists(paste(directory_output_reports, filename_rdata,sep=""))){first_analysis<-F}
if(first_analysis==T){
save(list = ls(all.names = TRUE), file = directory_and_filename_rdata, envir = environment())
}

```



```{r, child=if (2 %in% model_strategies_to_run) 'Ropls_with_permutated_variable_selection_Models_of_each_comparison_strategy2.Rmd'}

```

```{r}

if(first_analysis==T){
save(list = ls(all.names = TRUE), file = directory_and_filename_rdata, envir = environment())
}

```


```{r, child=if (3 %in% model_strategies_to_run) 'Ropls_with_permutated_variable_selection_Models_of_each_comparison_strategy3.Rmd'}

```

```{r}

if(first_analysis==T){
save(list = ls(all.names = TRUE), file = directory_and_filename_rdata, envir = environment())
}

```

```{r, child=if (4 %in% model_strategies_to_run) 'Ropls_with_permutated_variable_selection_Models_of_each_comparison_strategy4.Rmd'}

```

```{r}

if(first_analysis==T){
save(list = ls(all.names = TRUE), file = directory_and_filename_rdata, envir = environment())
}

```

```{r, child=if (5 %in% model_strategies_to_run) 'Ropls_with_permutated_variable_selection_Models_of_each_comparison_strategy5.Rmd'}

```

```{R}

save(list = ls(all.names = TRUE), file = directory_and_filename_rdata, envir = environment())

rm(list = setdiff(ls(envir = environment()), c("model_table_to_analyse", "filename_model_table_to_analyse", "directory_model_table_to_analyse", "projectname", "date_of_analysis", "directory_Rmarkdownfiles", "filename_Rmarkdownfile_each_model", "filename_Rmarkdownfile_summary", "directory_input_matrix_sampleID", "filename_matrix", "decimal_separator", "variable_names_length", "variable_names_position", "filename_sampleID", "colname_groupID", "colname_secID", "directory_output_reports", "directory_and_filename_function_file", "p_pearson_of_pcorr_cutoff", "no_permutations_post_vs", "no_permutations_post_vs_selected_models", "no_permutations_over_vs", "replace_0",  "filter_percent_in_each_group", "replace_NA","log_transform", "directory_output_reports_and_projectname", "groupsnumeric", "max_no_of_ortho_pre_vs", "max_no_of_ortho_post_vs", "each_model_or_summary", "prefered_pR2_and_pQ2_permutated_post_vs", "cluster", "reordered_levels_of_groups", "pcorr_diff", "variable_selection_using_VIP", "model_strategies_to_run")))
```
