---
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
mathjax: null

---

---
title: |
    | Model summary pre variable selection

author: |
    | roplspvs: Version 0.16.0
    | R orthogonal projections of latent structures with permutation over variable selection
    | Marika Strom
---

# Model strategy `r model_strategy`. `r model_description`


```{r, echo=FALSE}
summary_Model_sans_vs_table <-
  create_table_of_models_sans_vs(
    no_permutations_sans_vs,
    directory_output_reports,
    projectname,
    date_of_analysis,
    max_no_of_ortho_pre_vs
  )

summary_Model_sans_vs_table$`significant` <-
  (
    summary_Model_sans_vs_table$`pQ2 perm. sans v.s.` <= 0.05 &
      summary_Model_sans_vs_table$`R2Y(cum) sans v.s.` >= 0.5 &
      summary_Model_sans_vs_table$`Q2(cum) sans v.s.` >= 0.4
  )

summary_Model_sans_vs_tabletokable <- summary_Model_sans_vs_table
if (groupsnumeric == "yes") {
  order_of_models <-
    order(as.numeric(as.character(
      summary_Model_sans_vs_tabletokable$group1
    )), as.numeric(as.character(
      summary_Model_sans_vs_tabletokable$group2
    )))
  summary_Model_sans_vs_tabletokable <-
    summary_Model_sans_vs_tabletokable[order_of_models, ]
}
summary_Model_sans_vs_tabletokable[, c(6, 7, 8, 9, 12, 13, 14)] <-
  format(round(summary_Model_sans_vs_table[, c(6, 7, 8, 9, 12, 13, 14)], 3))
#summary_Model_sans_vs_tabletokable[,c(12,13)] <- format(summary_Model_sans_vs_table[,c(12,13)], scientific=TRUE, digits=2)

kable(summary_Model_sans_vs_tabletokable, row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
meanofR2 <- mean(summary_Model_sans_vs_table$`pR2Y perm. sans v.s.`)
meanofQ2 <- mean(summary_Model_sans_vs_table$`pQ2 perm. sans v.s.`)
meanofR2 <- format(meanofR2, digits = 2)
meanofQ2 <- format(meanofQ2, digits = 2)

```
Average of pR2Y respectively pQ2 of models with permutations sans v.s.:
pR2Y: `r meanofR2`
pQ2: `r meanofQ2`

```{r, echo=FALSE}

percentofsignmodel <- sum(summary_Model_sans_vs_tabletokable$significant)/nrow(summary_Model_sans_vs_tabletokable)*100
percentofsignmodel <- format(percentofsignmodel, digits=2)

```
`r percentofsignmodel`% of models were significant (p(Q2_perm_pre_vs) <= 0.05 and R2Y >= 0.5 and Q2 >= 0.4).

