---
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
mathjax: null

---

---
title: |
    | Model summary iteration strategy

author: |
    | roplspvs: Version 0.16.0
    | R orthogonal projections of latent structures with permutation over variable selection
    | Marika Strom
---

# Model strategy `r model_strategy`. `r model_description`



```{r, echo=FALSE}

summary_Model5_table <- create_table_of_models_iterations(directory_output_reports,projectname,date_of_analysis)
if (groupsnumeric == "yes"){
 summary_Model5_table <- summary_Model5_table[order_of_models,]
}
summary_Model5_tabletokable <- summary_Model5_table
summary_Model5_tabletokable[,c(9,10,11,12)] <- format(round(summary_Model5_tabletokable[,c(9,10,11,12)],3))

kable(summary_Model5_tabletokable, row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```



# Significant model strategy `r model_strategy`.  `r model_description` 
Showing significant models (p(Q2_perm_post_vs) <= 0.05 and R2Y >= 0.5 and Q2 >= 0.4)
Note that iteration models have not been permutated over variable selection


```{r, echo=FALSE}

summary_Model5_tablesignificant <- subset(summary_Model5_table,summary_Model5_table$`pQ2 permutated post v.s.`<=0.05 & summary_Model5_table$`R2Y(cum)`>=0.5 & summary_Model5_table$`Q2(cum)`>=0.4)
summary_Model5_tablesignificanttokable <- summary_Model5_tablesignificant

summary_Model5_tablesignificanttokable[,c(9,10,11,12)] <- format(round(summary_Model5_tablesignificanttokable[,c(9,10,11,12)],3))

kable(summary_Model5_tablesignificanttokable, row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

percentofsignmodel <- nrow(summary_Model5_tablesignificant)/nrow(summary_Model5_table)*100
percentofsignmodel <- format(percentofsignmodel, digits=0)



```
`r percentofsignmodel`% of models were significant.

